Comparing Facebook's Prophet Forecasting to an ARIMA Model: A Diebold-Mariano Evaluation of the Top40 Index

### 2017-10-17 template/Draft_Paper.Rmd
Added the DM Test mathematical formulation and moved section 2.1.4 to 2.4

Though the ARIMA model is tractable given that it is simple, interpretable and yields forecasts that are significantly accurate when compared to other methods, it has its limitations. The most popular being its inability to capture non-linear patterns in data, even after its evolution from the standard form to more adaptable formulations [@moreno2011artificial]. Over the past two to three decades with the evolution of computational power and statistical advancement, other stock price prediction methods have been proposed, the most prominent being a machine learning approach called Artificial Neural Networks(ANNs)[@lin2009short]. Also popular amongst the new stock price forecasting approaches are hybrids of existing methods that incorporate the benefits of different approaches.

@@

###Diebold-Mariano Evaluation
As is evident from the previous sections, the modelling approach and assumptions of Prophet differ to those of an ARIMA model. Since the two methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as a means of testing forecasting accuracy. This is because the DM Test is a model free test of forecasting accuracy. It is applicable to a wide range of situations from multi-period forecasts, non-Gaussian forecast residuals, non-quadratic loss functions and even serially correlated data[@Mariano2000].

@@

The challenge with many forecasting methods such as the Box-Jenkins methodology and machine-learning approaches are that they usually require a sufficient statistical background. Furthermore, these approaches tend to be inflexible with respect to incorporating prior information. Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates[@taylor2017forecasting].  
Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts. This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models. Prophet combines a configurable model that includes performance analysis evaluation with the interaction of an analyst in the loop [@taylor2017forecasting]. The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].

@@

>>>>

Though the ARIMA model is tractable given that it is simple, interpretable and yields forecasts that are significantly accurate when compared to other methods, it has its limitations. The most popular being its inability to capture non-linear patterns in data, even after its evolution from the standard form to more adaptable formulations [@moreno2011artificial]. Over the past two to three decades with the evolution of computational power and statistical advancement, other stock price prediction methods have been proposed, the most prominent being a machine learning approach called Artificial Neural Networks(ANNs) [@lin2009short]. Also popular amongst the new stock price forecasting approaches are hybrids of existing methods that incorporate the benefits of different approaches.

@@

@@

The challenge with many forecasting methods such as the Box-Jenkins methodology and machine-learning approaches are that they usually require a sufficient statistical background. Furthermore, these approaches tend to be inflexible with respect to incorporating prior information. Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates [@taylor2017forecasting].  
Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts. This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models. Prophet combines a configurable model that includes performance analysis evaluation with the interaction of an analyst in the loop [@taylor2017forecasting]. The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].

@@

##Diebold-Mariano Evaluation
As is evident from sections 2.2 and 2.3, the modelling approach and assumptions of Prophet differ to those of an ARIMA model. Since the two methods differ, the Diebold-Mariano evaluation is preferred as a means of comparing the forecast accuracy of the two different methods. The Diebold-Mariano evaluation is a **model free** hypothesis test for the equal predictive accuracy of two forecasting methods. It is applicable to a wide range of situations from forecast residuals that are non-Gaussian, asymmetric loss functions, serially correlated errors and multi-period forecasts [@Mariano2000]. The Diebold-Mariano test (DM test) operates by evaluating the forecast errors for both model 1 and model 2 (i.e $e_{t_1}$ and $e_{t_2}$) at a loss function $g(\bullet)$. The loss function can be choosen flexibly as long as it is zero when the error is zero, is stricly positive for all values on the real line and, it increases with an increase in the absolute value of the error size. The loss differential between two forecasts at time $t$ is then defined by, $$ d_t =  g(e_{t1}) - g(e_{t2})$$ for $t = (1, 2, ..., M)$ where $M$ is the total number of forecasts that have been generated from each model. We then test whether, $$ \mu = E[d_t] = 0 $$ by conducting the following hypothesis test;  
Null: $$ E[d_t] = 0  \,\,\,\, \forall t$$   
Alternative : $$ E[d_t] \neq 0 $$  
Test statistic: $$DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} $$

$\bar{d}$ is the sample mean of the $M$ loss differentials and $\frac{\hat{\gamma_d(0)}}{M}$ is the asymptotic variance of  ($\bar{d} - \mu$) with $\hat{\gamma_d(k)}$ representing the autocovariance of the loss differential series at lag $k=0$. Asymptotically, the $DM_{stat}$ has a standard normal distribution if the loss differential series is short memory and covariance stationary [@Mariano2000].  
There are two caveats to using the DM Test for the purposes of this paper. Firstly,both a linear and quadratic loss function was  <<<<<<**correct the tense!!!!!**>>>>> used to conduct the DM Test since literature does not clearly stipulate which method is more reliable **CITE???**. Using either a linear or quadratic loss function implicitly assumes that the forecast errors are symmetric. The assumption  which is a reasonable assumption since **<<<<<<JUSTIFY THE ASSUMPTION BY REFERENCING THE BOXPLOT WHERE ERRORS LOOKED SYMMETRIC >>>>>>>**. secondly, literature does not clearly stipulate what sample size is sufficient for the asymptotic distribution assumption to hold. In response to this, 20% of the available data was used as the out-of-sample period. Nonethless, the DM test results must used conservatively .

### 2017-10-17 template/Draft_Paper.tex
Added the DM Test mathematical formulation and moved section 2.1.4 to 2.4

Though the ARIMA model is tractable given that it is simple,
interpretable and yields forecasts that are significantly accurate when
compared to other methods, it has its limitations. The most popular
being its inability to capture non-linear patterns in data, even after
its evolution from the standard form to more adaptable formulations
(Moreno, Pol, and Gracia 2011). Over the past two to three decades with
the evolution of computational power and statistical advancement, other
stock price prediction methods have been proposed, the most prominent
being a machine learning approach called Artificial Neural
Networks(ANNs)(Lin, Yang, and Song 2009). Also popular amongst the new
stock price forecasting approaches are hybrids of existing methods that
incorporate the benefits of different approaches.

@@

\subsubsection{Diebold-Mariano
Evaluation}\label{diebold-mariano-evaluation}

As is evident from the previous sections, the modelling approach and
assumptions of Prophet differ to those of an ARIMA model. Since the two
methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as
a means of testing forecasting accuracy. This is because the DM Test is
a model free test of forecasting accuracy. It is applicable to a wide
range of situations from multi-period forecasts, non-Gaussian forecast
residuals, non-quadratic loss functions and even serially correlated
data(Mariano 2000).

@@

The challenge with many forecasting methods such as the Box-Jenkins
methodology and machine-learning approaches are that they usually
require a sufficient statistical background. Furthermore, these
approaches tend to be inflexible with respect to incorporating prior
information. Though this can be partially achieved by initializing
parameter estimates for some methods (such as ANNs) ,an analyst with no
statistical training can have difficulty interpreting and thus modifying
parameter estimates(Taylor and Letham 2017).\\
Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
Generalized Additive model to generate time series forecasts. This is in
comparison to conventional methods such as the linear stochastic
dependence that exists in ARIMA models. Prophet combines a configurable
model that includes performance analysis evaluation with the interaction
of an analyst in the loop (Taylor and Letham 2017). The model is
configurable which allows an analyst to incorporate their knowledge of
the behaviour of the series into the model building process through
easily interpretable initial parameters that can be modified and
interactive feedback when forecasts under-perform (Taylor and Letham
2017).

@@

>>>>

Though the ARIMA model is tractable given that it is simple,
interpretable and yields forecasts that are significantly accurate when
compared to other methods, it has its limitations. The most popular
being its inability to capture non-linear patterns in data, even after
its evolution from the standard form to more adaptable formulations
(Moreno, Pol, and Gracia 2011). Over the past two to three decades with
the evolution of computational power and statistical advancement, other
stock price prediction methods have been proposed, the most prominent
being a machine learning approach called Artificial Neural
Networks(ANNs) (Lin, Yang, and Song 2009). Also popular amongst the new
stock price forecasting approaches are hybrids of existing methods that
incorporate the benefits of different approaches.

@@

@@

The challenge with many forecasting methods such as the Box-Jenkins
methodology and machine-learning approaches are that they usually
require a sufficient statistical background. Furthermore, these
approaches tend to be inflexible with respect to incorporating prior
information. Though this can be partially achieved by initializing
parameter estimates for some methods (such as ANNs) ,an analyst with no
statistical training can have difficulty interpreting and thus modifying
parameter estimates (Taylor and Letham 2017).\\
Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
Generalized Additive model to generate time series forecasts. This is in
comparison to conventional methods such as the linear stochastic
dependence that exists in ARIMA models. Prophet combines a configurable
model that includes performance analysis evaluation with the interaction
of an analyst in the loop (Taylor and Letham 2017). The model is
configurable which allows an analyst to incorporate their knowledge of
the behaviour of the series into the model building process through
easily interpretable initial parameters that can be modified and
interactive feedback when forecasts under-perform (Taylor and Letham
2017).

@@

\subsection{Diebold-Mariano
Evaluation}\label{diebold-mariano-evaluation}

As is evident from sections 2.2 and 2.3, the modelling approach and
assumptions of Prophet differ to those of an ARIMA model. Since the two
methods differ, the Diebold-Mariano evaluation is preferred as a means
of comparing the forecast accuracy of the two different methods. The
Diebold-Mariano evaluation is a \textbf{model free} hypothesis test for
the equal predictive accuracy of two forecasting methods. It is
applicable to a wide range of situations from forecast residuals that
are non-Gaussian, asymmetric loss functions, serially correlated errors
and multi-period forecasts (Mariano 2000). The Diebold-Mariano test (DM
test) operates by evaluating the forecast errors for both model 1 and
model 2 (i.e \(e_{t_1}\) and \(e_{t_2}\)) at a loss function
\(g(\bullet)\). The loss function can be choosen flexibly as long as it
is zero when the error is zero, is stricly positive for all values on
the real line and it increases with an increase in the absolute value of
the error size. The loss differential between two forecasts at time
\(t\) is then defined by, \[ d_t =  g(e_{t1}) - g(e_{t2})\] for
\(t = (1, 2, ..., M)\) where \(M\) is the total number of forecasts that
have been generated from each model. We then test whethe,r
\[ \mu = E[d_t] = 0 \] under the following hypothesis test;\\
Null: \[ E[d_t] = 0  \,\,\,\, \forall t\]\\
Alternative : \[ E[d_t] \neq 0 \]\\
Test statistic:
\[DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} \]

\(\bar{d}\) is the sample mean of the \(M\) loss differentials and
\(\frac{\hat{\gamma_d(0)}}{M}\) is the asymptotic variance of
(\(\bar{d} - \mu\)) with \(\hat{\gamma_d(k)}\) representing the
autocovariance of the loss differential series at lag \(k=0\). The
asymptotic distribution of \(DM_{stat}\) is then standard normal
asumming the loss differential series is short memory and covariance
stationary (Mariano 2000).\\
There are two caveats to using the DM Test for the puposes of this
paper. Firstly,both a linear and quadratic loss function was
\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}\textbf{correct
the
tense!!!!!}\textgreater{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}
used to conduct the DM Test since literature does not clearly stipulate
which method is more reliable \textbf{CITE???}. Using either a linear or
quadratic loss function implicitly assumes that the forecast errors are
symmetric. The assumption which is a reasonable assumption since
\textbf{\textless{}\textless{}\textless{}\textless{}\textless{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}}.
secondly, literature does not clearly stipulate what sample size is
sufficient for the asymptotic distribution assumption to hold. In
response to this, 20\% of the available data was used as the
out-of-sample period. Nonethless, the DM test results must used
conservatively .
